---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---
<br/>

Hi! Now we are going the start this assignment choosing a library for our plots.

<br/>

**0. Choosing library plot**
<br/>
```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```

----

### Loading and preprocessing the data
<br/>

**1. We need to unzip the file "activity.zip":**

```{r}
temp<-"C:/Users/Leticia/Desktop/RepData_PeerAssessment1/activity.zip"
unzip(temp)
```
<br/>

**2. Now we can read the csv file using read.csv:**

```{r}
activity<-read.csv("activity.csv")
```


```{r}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```

<br/>

Ok! So, let's do some data analysis! 

----

### What is mean total number of steps taken per day?

<br/>

**3. Computes the number of steps per day in *stepsByDay* variable:**


```{r}
stepsByDay <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```

<br/>

**4. Make an histogram:**


```{r}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

<br/>

**5. Calculate and report the mean and median total number of steps taken per day.**


```{r}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```


* Mean: `r stepsByDayMean`
* Median:  `r stepsByDayMedian`

----


### What is the average daily activity pattern?

**6. xxx**
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)
```

<br/>
**7. xxx**
```{r}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```
<br/>

----

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

**8. xxx**

```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```

<br/>

* Most Steps at: `r timeMostSteps`

----

### Imputing missing values

<br/>
**9. xxx**
```{r}
numMissingValues <- length(which(is.na(activity$steps)))
```
<br/>

* Number of missing values: `r numMissingValues`

<br/>

**10. Devise a strategy for filling in all of the missing values in the dataset.**

<br/>

**11. Create a new dataset that is equal to the original dataset but with the missing data filled in.**
```{r}
activityDataImputed <- activity
activityDataImputed$steps <- impute(activity$steps, fun=mean)
```

<br/>

----

### Are there differences in activity patterns between weekdays and weekends?



**12. Make a histogram of the total number of steps taken each day.**
```{r}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

<br/>

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
* Mean (Imputed): `r stepsByDayMeanImputed`
* Median (Imputed):  `r stepsByDayMedianImputed`


----

### Are there differences in activity patterns between weekdays and weekends?

<br/>
**1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.**

```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

**2. Make a panel plot containing a time series plot.**

```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```
